Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating

### 2017-08-25 R/SubGenome Analysis.Rmd
Added some analysis for expression data

---
title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
  subgenome assignments in maize'
params:
  fileParser: ../Perl/parseKsKnFile.pl
  goDataFile: ../Data/GO from maizecyc.tab
  log10_ks_cutoff: 0
  # synMapOutputFile: 6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks
  # inputDataFile: ../Data/sorghum_v1_vs_maize_v1.tab
  # subGenomeFile: ../Data/outfile_processed_byHand.tab

  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
  subGenomeFile: ../Data/outfile_processed_byHand.tab
output:
  html_document: default
  pdf_document: default
---

@@

```{r importSynMapData, include=FALSE}
source("~/git/SubGenomes/R/parseSubgenomes.R")
source("~/git/SubGenomes/R/addGOTerms.R")
source("~/git/SubGenomes/R/parseGOSlim.R")
# source("~/git/SubGenomes/R/createTopGO.R")

MFterms <- ls(GOMFTerm)

@@

***  
 * There were **`r syntelogs.raw %>% nrow()`** rows of data in the SynMap file.
 * There were **`r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
 * Summary of the KS values
     + **`r syntelogs.raw %>% select(ks) %>% summary()`**
 * Out of the `r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
 * SubGenome 1
     + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
 * SubGenome 2
     + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
 * Rejected
     + **`r subgenome %>% filter(subgenome == "rejected:small") %>% select(gene2) %>% distinct() %>% nrow()`** genes were rejected during subgenome parsing
     (Too few syntenic genes across chromosome)

@@

```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2
converted <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
converted <- 
  converted %>%
  select(Maize2, "GO Term", EVCode) %>%
  rename(gene2=Maize2)

@@

>>>>

---
title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
  subgenome assignments in maize'
params:
  fileParser: ../Perl/parseKsKnFile.pl
  goDataFile: ../Data/GO from maizecyc.tab
  inputDataFile: ../Data/sorghum_v3.1_vs_maize_v4+rejected.tab
  log10_ks_cutoff: 0
  subGenomeFile: ../Data/outfile_processed_byHand.tab
output:
  html_document: default
  pdf_document: default
---

@@

```{r importSynMapData, include=FALSE}
source("~/git/SubGenomes/R/parseSubgenomes.R")
source("~/git/SubGenomes/R/addGOTerms.R")
source("~/git/SubGenomes/R/parseGOSlim.R")
source("~/git/SubGenomes/R/createTopGO.R")
source("~/git/SubGenomes/R/parseExpressionData.R")
MFterms <- ls(GOMFTerm)

@@

***  
 * There were **`r syntelogs.raw %>% nrow()`** rows of data in the SynMap file.
 * There were **`r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()`** distinct Maize genes with synteny to Sorghum.
 * Summary of the KS values
     + **`r syntelogs.raw %>% select(ks) %>% summary()`**
 * Out of the `r syntelogs.raw %>% select(gene2) %>% distinct() %>% nrow()` distinct Maize genes with synteny to sorghum, we keep **`r homeologs.genes %>% select(gene2) %>% distinct() %>% nrow()`** after applying the KS cutoff.
 * SubGenome 1
     + **`r subgenome %>% filter(subgenome == "sub1") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 1
     + **`r subgenome %>% filter(subgenome == "sub1") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 1 that are expressed significantly on average over 68 samples
     + **`r goAnnotations.sub1 %>% nrow()`** GO assignments using **`r goAnnotations.sub1 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
 * SubGenome 2
     + **`r subgenome %>% filter(subgenome == "sub2") %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned to subgenome 2
     + **`r subgenome %>% filter(subgenome == "sub2") %>% filter(gene2 %in% expressedGenes$geneID) %>% select(gene2) %>% distinct() %>% nrow()`** genes assigned            to subgenome 2 that are expressed significantly on average over 68 samples
     + **`r goAnnotations.sub2 %>% nrow()`** GO assignments using **`r goAnnotations.sub2 %>% select('GO Term') %>% distinct() %>% nrow()`** GO Terms
 * Rejected
     + **`r subgenome %>% filter(subgenome == "rejected:small") %>% select(gene2) %>% distinct() %>% nrow()`** genes were rejected during subgenome parsing
     (Too few syntenic genes across chromosome)

@@

```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
converted <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
converted <- 
  converted %>%
  select(Maize2, "GO Term", EVCode) %>%
  rename(gene2=Maize2)

@@

# Expression

Cases where the maize1 homolog is more or less expressed than the maize2 homolog

```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
## Attach log2(FPKM_mean) values to homeologous pairs
expressedPairs <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  inner_join(expressedGenes, by=c("Maize1"="geneID")) %>%
  rename(FPKM_mean1=FPKM_mean) %>%
  inner_join(expressedGenes, by=c("Maize2"="geneID")) %>%
  rename(FPKM_mean2=FPKM_mean)

# Expression Diffs
grid.newpage()
g <- draw.pairwise.venn(
  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
                          nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
```
